---
title: "LandUse+ extension"
author: "Núria Aquilué"
date: "9/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(viridis)
```
The objective is to set-up the LANDIS-II extension LandUse+ to simulate land-cover changes.


## Dummy example

I build a 4x4 landscape.  
The potential land-covers are:  
## Dummy example
* urban areas (red)
* forest (violet)
* croplands (yellow)
* oaks plantations (green)
* pines plantations (blue).  

The initial landscape (at year 0) only has urban areas and forest.  
The next years (y10 and y20) crops appear, in y30 nothing changes.  
At year 40 oaks are planted in previous forest land.  
At year 60 pines should be planted in urban areas.  

```{r land-cover-map, echo=FALSE}
setwd("C:/WORK/FUNCT.NET/TOY")
par(mfrow=c(2,4))
land <- raster("Model/inputlyrs/land_cover0.img")
plot(land, col=rainbow(5), main="Land-cover y0", legend=F)
land <- raster("Model/inputlyrs/land_cover10.img")
plot(land, col=rainbow(5), main="Land-cover y10", legend=F)
land <- raster("Model/inputlyrs/land_cover20.img")
plot(land, col=rainbow(5), main="Land-cover y20", legend=F)
plot(land, col=rainbow(5), main="Land-cover y30", legend=F)
land <- raster("Model/inputlyrs/land_cover40.img")
plot(land, col=rainbow(5), main="Land-cover y40", legend=F)
plot(land, col=rainbow(5), main="Land-cover y50", legend=F)
land <- raster("Model/inputlyrs/land_cover60.img")
plot(land, col=rainbow(5), main="Land-cover y60", legend=F)
plot(land, col=rainbow(5), main="Land-cover y70", legend=F)
```

At year 0, up to 7 types of forest communities are randomly allocated in forest area and 3 ecoregions are initialized.  
**In urban areas (at year 0), I assign 0 to initial forest communities and ecoregion**
```{r forest-map, echo=FALSE}
setwd("C:/WORK/FUNCT.NET/TOY")
par(mfrow=c(1,2))
forest <- raster("Model/inputlyrs/initial_communities.img")
plot(forest, col=c("grey", rainbow(7)), main="Initial tree communities")
eco <- raster("Model/inputlyrs/ecoregions.img")
plot(eco, col=c("grey", viridis(3)), main="Ecoregions")
```

I run LANDIS-II with "Land Use Change" extension specified in *land-use.txt* (see attached).  

**The biomass maps look like:**    
```{r biomass-map, echo=FALSE}
setwd("C:/WORK/FUNCT.NET/TOY")
## Biomass maps
species <- c("abiebals", "acerrubr", "acersacc", "betualle", "betupapy", "fraxamer", "piceglau", "pinubank", 
             "pinuresi", "pinustro", "poputrem", "querelli", "querrubr", "thujocci", "tiliamer", "tsugcana")
par(mfrow=c(2,4))
for(year in seq(0,70,10)){
  biom <- raster(paste0("Model/outputs/biomass/abiebals-",year,".img"))
  for(spp in species[-1]){
    aux <- raster(paste0("Model/outputs/biomass/",spp,"-",year,".img"));  biom <- biom + aux
  }
  # print(paste0("Biomass y", year, " per pixel: "), biom[]/1000))
  plot(biom/1000, col=c("grey", viridis(4)), main=paste0("Biomass y", year))               
}
```

### Conversion from forest to crops WORKS

### Conversion from previous forest, to crop, then to plantations WORKS

### Conversion from non-forest to forest DOES NOT WORK
I want to convert non-forest zones at the beginning of the simulation to forest in the future (y60).   
These non-forest zones at year 0 do not need any cohort assigned.  
I assign MapCode 0, and MapCode 0 doesn't have any cohort, so anything grows 
(even if I don't eventually a cohort may establish there).  
The second option is to assign to these non-forest zones the Ecoregion 0 and this be innactive.  
In any case, I cannot turn non-forest to forest.
